Isomorphic Interpreters from Logically Reversible Abstract Machines

### 2012-09-14 iso-int/rc-iso-int.tex
_

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parity Translation }
\label{sec:parity-int}

While the previous machines expressed traversals of the values with
input values reflected in the outputs, the following machine
translates one recursive type to another. Consider the type {{nat}}
(defined to be {{@@x.1+x}}) and its parity encoded version {{parity}}
(defined to be {{@@x.1+(1+x)}}). Algebraically, we may write:

%subcode{bnf} include main
% Numbers, n, m = 0 | n + 1
% Parity, par = even | odd | A par

While it is apparent that there is a natural one to one correspondence
between their values, it is non-obvious how this can be expressed as a
{{langRevT}} isomorphism. This map can be expressed in terms of the
algebraic type definitions as shown below:

%subcode{opsem} include main
%! columnStyle = ||c|r||
% 0 & even
% 1 & odd
% 2 & A even
% 3 & A odd
% 4 & A (A even)
% ... &  ...

\indent
and in terms of their {{langRevT}} type encodings as shown below:

%subcode{opsem} include main
%! columnStyle = ||c|c||
% <left ()> & <left ()>
% <right <left ()>> & <right left ()>
% <right <right <left ()> > > & <right right <left ()>>
% <right <right <right <left ()>> >> & <right right <right left ()>>
% <right <right <right <right <left ()>> >> > & <right right <right right <left ()> > >
% ... &  ...

\noindent
Instead of directly programming in {{langRevT}}, one can derive a
{{langRevT}} combinator for this purpose by first constructing an
abstract machine that maps {{nat}} to {{parity}}.  A logically
reversible abstract machine expressing the translation may be defined
as follows:

%subcode{bnf} include main
% Machine states = <nat, nat> | {[parity, nat]}
% Start state = <n , 0>
% Stop state =  {[par, 0]}

\noindent
Machine Reductions:
%subcode{opsem} include main
%! columnStyle = rclr
% <0, m> &|-->& {[even, m]}
% <1, m> &|-->& {[odd, m]}
% <(n+1)+1, m> &|-->& <n, m+1>
% {[par, m+1]} &|-->& {[A par, m]}

The construction of the {{langRevT}} combinator is straightforward and
proceeds as before. The partial combinator representing the wiring of
machine reductions is shown below.

\begin{center}
\scalebox{1.0}{
  \includegraphics{diagrams/nat-parity.pdf}
}
\end{center}

\noindent
We thank Fritz Henglein for the motivation underlying this example.
